When does vapor pressure deficit drive or reduce evapotranspiration?

### 2017-12-15 doc/agu_2017/vpd.tex
first commit of messy

\section{Method}
\begin{frame}{Develop a theory to quantify VPD effects}
  \begin{itemize}
  \item The goal is to simply the problem (increase transparency) while still capturing the leading order behavior the system.
    \begin{itemize}
    \item Simplifying the system aids intrinsic understanding, but at a physical realism.
    \end{itemize}
  \item Because we run the risk of over-simplification, we will use FLUXNET2015 data to test how well our theory matches the data.
    \begin{itemize}
    \item We will test in the growing season of 5 plant types: Deciduous broadleaf forest, evergreen needleleaf forest, shrub, grass, and crops.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Test theory with FLUXNET data - the good}
  \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.8\textheight)
    \textcolor{blue}{Blue line} is theory's critical VPD.\\
  \end{textblock*}

  \begin{columns}
    \begin{column}{0.33\textwidth}<1-2>
      \includegraphics[width=1.1\textwidth]{DBF_box.pdf}
    \end{column}
    \begin{column}{0.33\textwidth}<2>
      \includegraphics[width=1.1\textwidth]{ENF_box.pdf}
    \end{column}
    \begin{column}{0.33\textwidth}<2>
      \includegraphics[width=1.1\textwidth]{CSH_box.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Test theory with FLUXNET data - the bad}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{GRA_box.pdf}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{CRO_box.pdf}
    \end{column}
  \end{columns}
  \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.85\textheight)
    \textcolor{blue}{Blue line} is theory's critical VPD.\\
  \end{textblock*}

@@

@@

>>>>

\section{Method}
\begin{frame}{Develop a theory to quantify ET response to VPD}
  \begin{itemize}
  \item The goal is to simply the problem (increase transparency) while still capturing the leading order behavior the system.
    \begin{itemize}
    \item Simplifying the system aids intrinsic understanding, but at a physical realism.
    \end{itemize}
  \item Because we run the risk of over-simplification, we will use FLUXNET2015 data to test how well our theory matches the data.
    \begin{itemize}
    \item We will test in the growing season of 5 plant types: deciduous broadleaf forest, evergreen needleleaf forest, shrub, grass, and crops.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Test theory with FLUXNET data - the good}
  \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.8\textheight)
    \textcolor{blue}{Blue line} is theory.\\
  \end{textblock*}

  \begin{columns}
    \begin{column}{0.33\textwidth}<1-3>
      \includegraphics[width=1.1\textwidth]{DBF_box.pdf}
    \end{column}
    \begin{column}{0.33\textwidth}<2-3>
      \includegraphics[width=1.1\textwidth]{CSH_box.pdf}
    \end{column}
    \begin{column}{0.33\textwidth}<3>
      \includegraphics[width=1.1\textwidth]{ENF_box.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Test theory with FLUXNET data - the bad}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{GRA_box.pdf}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{CRO_box.pdf}
    \end{column}
  \end{columns}
  \begin{textblock*}{1.0\textwidth}(0.3\textwidth,0.85\textheight)
    \textcolor{blue}{Blue line} is theory.\\
  \end{textblock*}

@@

\begin{frame}{Summary - When does VPD drive or reduce ET?}
  \begin{itemize}
  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmospheric demand dominates).
  \item For forest sites, environmental VPD approximately straddles the critical VPD.
  \item For shrubs environmental VPD never exceeds the critical VPD.
  \item Theory tested poorly with FLUXNET data for for crops and grass. 
  \item All plant types exhibited a response far below that of PET.
  \item The new uWUE-version of Penman-Monteith we derived could be used as a complement for PET in drought indices over vegetated surfaces.
  \end{itemize}
  \bigskip
  \medskip
  Adam Massmann:\textbf{ akm2203@columbia.edu }
\end{frame}

@@

\begin{frame}{Consequences of theory - ``sign'' term}
Below is a little opaque (at least to me):
\[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
so do expansion about $g_1 >> \sqrt{VPD}$:
\footnotesize
\[ \frac{\partial \, ET}{\partial \, VPD} \approx \text{scaling} \cdot  \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{1}{g_1} - \frac{3\sqrt{VPD}}{2 g_1^2} +  \mathcal{O}\left(\frac{\sqrt{VPD}}{g_1}\right)^{2} \right) \right)\]
% \large
% and:
%   \[ \sqrt{VPD} \Big|_{\frac{\partial ET}{\partial VPD} = 0} \approx  \frac{2 g_1}{3}  \left( -\frac{ c_p \, g_1\,  1.6  \; R*\; \text{ uWUE }}{R_{air} \gamma c_s } + 1  \right) \]

%   \[ VPD \Big|_{\frac{\partial ET}{\partial VPD} = 0} = \frac{R_{air}}{4 \, c_p} \left[ \frac{\gamma  c_s}{R* \, 1.6\, \text{uWUE}} + \sqrt{\left(g_1 + 8\, \frac{c_p}{R_{air}}\right) \frac{\gamma c_s}{R* \, 1.6\, \text{uWUE}}} - 4 \, g_1 \frac{ c_p}{R_{air}}\right] \]
\end{frame}
